---
title: "Network Graph"
author: "Elizabeth Nieto"
date: "November 20, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#REFERENCES :
https://www.jessesadler.com/post/network-analysis-with-r/
https://www.rmanchester.org/wp-content/uploads/sites/3/presentations/ManchesterR_-_Network_Visualisation_with_ggraph_-_Hannah_Frick_-_20180206.pdf
https://www.data-imaginist.com/2017/announcing-ggraph/
http://mr.schochastics.net/netVizR.html
https://cran.r-project.org/web/packages/ggraph/ggraph.pdf

#REFERENCES FOR FUTURE WORK w/ different network packages in R 
https://briatte.github.io/ggnet/
```{r}

library(tidygraph)
library(ggraph)
library(tidyr)
library(plyr)
library(visNetwork)
library(networkD3)


#set working directory
setwd("C:\\Users\\eniet\\Documents\\School\\Autumn19\\DSC465_Data_Visualization\\Project")

getwd()
```
#GET data
#GET sport data
#GET medal count data
```{r}

data <-read.csv("olympics.csv")
head(data,15)

#cast(data,ID ~ Event)
#filter out athletes w/o medals
winners <- data[!is.na(data$Medal),]
#filter out all sports except Alpine Skiing
alpine_winners <- filter (winners, Sport == 'Alpine Skiing')
#filterout any case that was played before 1988 bc this is when all events started
alpine_winners_1988_2016 <- filter(alpine_winners, Year > 1987)
#select data set with only ID and events
id_events <- select(alpine_winners_1988_2016 ,ID,Year,Medal,Event)
#241 Rows
#this data sets has one row contianing id and event per athlete who won in alpine skiing
id_events
#filter out athletes who only won one medal
athleteIDs <- select(id_events, ID)
ath_freq <- athleteIDs %>% 
  count('ID')
multiple_wins <- filter(ath_freq, freq != '1')
#filter to only get unique events per id
#id_events %>%
#  distinct() 
#join athletes w/ multiple wins to all athletes
multiplewins <- join(multiple_wins,id_events, by = 'ID', type = 'inner')
#convert event values to numeric ids of alpine tibble evetn ids
#ABANDONED - replace(multiplewins$Event, multiplewins$Event == 'Alpine Skiing Men\'s Downhill', 1)

#pivot dataset to get each event to become a column containing a medal value
piv_event <- multiplewins %>%
  spread(Event, Medal)

#convert event values to numeric ids of alpine tibble evetn ids
renamed_events <- rename(piv_event, c("Alpine Skiing Men\'s Downhill" = '1', 
                    "Alpine Skiing Men's Super G" = '2',
                    "Alpine Skiing Men's Giant Slalom" = '3',
                    "Alpine Skiing Men's Slalom" = '4',
                    "Alpine Skiing Men's Combined" = '5',
                    "Alpine Skiing Women's Slalom" = '6',
                    "Alpine Skiing Women's Giant Slalom" = '7',
                    "Alpine Skiing Women's Super G" = '8',
                    "Alpine Skiing Women's Combined" = '9',
                    "Alpine Skiing Women's Downhill" = '10'
                    ))

```


#GET event data in tibble format
```{r}
#MENS EVENTS
#athletes who won in sport 1 and 2
events_1_2 <-filter(renamed_events,!((is.na(`1`)| is.na(`2`))))
line_1_2 <- nrow(events_1_2)

#athletes who won in sport 1 and 3
events_1_3 <- filter(renamed_events,!((is.na(`1`)| is.na(`3`))))
line_1_3 <- nrow(events_1_3)

#athletes who won in sport 1 and 4
events_1_4 <- filter(renamed_events,!((is.na(`1`)| is.na(`4`))))
line_1_4 <- nrow(events_1_4)

#athletes who won in sport 1 and 5
events_1_5 <- filter(renamed_events,!((is.na(`1`)| is.na(`5`))))
line_1_5 <- nrow(events_1_5)

#athletes who won in sport 2 and 3
events_2_3 <- filter(renamed_events, !((is.na(`2`)| is.na(`3`))))
line_2_3 <- nrow(events_2_3)

#athletes who won in sport 2 and 4
events_2_4 <- filter(renamed_events,!((is.na(`2`)| is.na(`4`))))
line_2_4 <- nrow(events_2_4)

#athletes who won in sport 2 and 5
events_2_5 <- filter(renamed_events,!((is.na(`2`)| is.na(`5`))))
line_2_5 <- nrow(events_2_5)

#athletes who won in sport 3 and 4
events_3_4 <- filter(renamed_events,!((is.na(`3`)| is.na(`4`))))
line_3_4 <- nrow(events_3_4)

#athletes who won in sport 3 and 5
events_3_5 <- filter(renamed_events,!((is.na(`3`)| is.na(`5`))))
line_3_5 <- nrow(events_3_5)

#athletes who won in sport 4 and 5
events_4_5 <- filter(renamed_events,!((is.na(`4`)| is.na(`5`))))
line_4_5 <- nrow(events_4_5)


#WOMENS EVENTS
#athletes who won in sport 6 and 7
events_6_7 <-filter(renamed_events,!((is.na(`6`)| is.na(`7`))))
line_6_7 <- nrow(events_6_7)

#athletes who won in sport 6 and 8
events_6_8 <- filter(renamed_events,!((is.na(`6`)| is.na(`8`))))
line_6_8 <- nrow(events_6_8)

#athletes who won in sport 6 and 9
events_6_9 <- filter(renamed_events,!((is.na(`6`)| is.na(`9`))))
line_6_9 <- nrow(events_6_9)

#athletes who won in sport 6 and 10
events_6_10 <- filter(renamed_events,!((is.na(`6`)| is.na(`10`))))
line_6_10 <- nrow(events_6_10)

#athletes who won in sport 7 and 8
events_7_8 <- filter(renamed_events,!((is.na(`7`)| is.na(`8`))))
line_7_8 <- nrow(events_7_8)

#athletes who won in sport 7 and 9
events_7_9 <- filter(renamed_events,!((is.na(`7`)| is.na(`9`))))
line_7_9 <- nrow(events_7_9)

#athletes who won in sport 7 and 10
events_7_10 <- filter(renamed_events,!((is.na(`7`)| is.na(`10`))))
line_7_10 <- nrow(events_7_10)

#athletes who won in sport 8 and 9
events_8_9 <- filter(renamed_events,!((is.na(`8`)| is.na(`9`))))
line_8_9 <-nrow(events_8_9)

#athletes who won in sport 8 and 10
events_8_10 <- filter(renamed_events,!((is.na(`8`)| is.na(`10`))))
line_8_9 <- nrow(events_8_10)

#athletes who won in sport 9 and 10
events_9_10 <- filter(renamed_events,!((is.na(`9`)| is.na(`10`))))
line_9_10<- nrow(events_9_10)

#backup filter(renamed_events,`1` == 'Gold')
```

#CREATE NEW DF for events
```{r}
from <- c(1,1,1,2,2,3,3,4,6,6,6,6,7,7,7,8,8,9)
to <- c(2,3,5,3,5,4,5,5,7,8,9,10,8,9,10,9,10,10)
weight <- c(7,4,4,7,3,3,2,3,6,1,7,1,4,5,3,4,4,5)

length(weight)
ec_df <- data.frame(from, to, weight)
names(ec_df) <- c('from', 'to', 'weight')


#event_counts <- rbind(event_counts, c(1,2,7), c(1,3,4), c(1,4,0), c(1,5, 4 ), c(2,3,7 ),  c(2,4, 0), c(2,5,3), c(3,4,3), c(3,5,2), c(4,5,3), c(6,7,6), c(6,8,1), c(6,9,7), c(6,10,1), c(7,8,4), c(7,9,5), c(7,10,3), c(8,9,4), #c(8,10,4), c(9,10,5) )
```
#TEST TO COMBINE NETWORKS - Create a new link list to merge event ids to only include 1-5 and add a gender column
```{r}
from <- c(1,1,1,2,2,3,3,4,1,1,1,1,2,2,2,3,3,4)
to <- c(2,3,5,3,5,4,5,5,2,3,4,5,3,4,5,4,5,5)
weight <- c(7,4,4,7,3,3,2,3,6,1,7,1,4,5,3,4,4,5)
gender <- c('m', 'm', 'm', 'm', 'm','m','m','m', 'f', 'f', 'f', 'f', 'f','f','f','f','f','f')

ec_df2 <- data.frame(from, to, weight, gender)
names(ec_df2) <- c('from', 'to', 'weight','gender')
```


#GET node data - alpine events
```{r}
events <- select(data, Event)
alpine <- filter (data, Sport == 'Alpine Skiing')
alpineevents <- unique(alpine$Event)
alpineevents
```
#TEST TO COMBINE NETWORKS - 
```{r}
fiveevents <- c('Downhill', 'Super G', 'Giant Slalom', 'Slalom','Combined' )
fiveevent_ids <- c(1,2,3,4,5)

ec_df2 <- data.frame(fiveevent_ids,fiveevents)
names(ec_df2) <- c('ID', 'event')
```

#GET event tibble 
#add ID column 
#change value col name to event
```{r}
alpineevents <- as_tibble(alpineevents)
alpineevents <- tibble::rowid_to_column(alpineevents,'ID')
#alpineevents <- alpineevents%>% 
#  rename('Event' = value )
alpineevents
class(alpineevents)
```
#GET event count tibble
```{r}
event_counts_tibble <- as_tibble(ec_df)

#event_counts_tibble <- plyr::rename(event_counts_tibble, c('Node1' = 'from', 'Node2' = 'to', 'Freq' = 'weight'))

```
#CREATE tidygraph
```{r}
events_tidy <- tbl_graph(nodes = alpineevents, edges = event_counts_tibble, directed = TRUE)

class(events_tidy)
```

#CREATE nETWORK GRAPH
```{r}
library (ggrepel)

ggraph(events_tidy, layout = "auto") +
  geom_node_point(color = 'blue') +
  geom_node_label(aes(label = value), color = 'blue', repel = TRUE)+
  
  geom_edge_link(aes(width = weight), alpha = .08) +
  
  scale_edge_width(range =c(0.2,2)) +
  labs(edge_width = 'Wins in Both') +
  theme_graph(background = 'white') +
  ggtitle('Alpine Skiing')


```
#CREATE INTERACTIVE Network Graph
```{r}

net<- visNetwork(alpineevents, ec_df)
plot(net)

